null

### 2017-04-03 Manuscript/Manuscript_COD_Trends.Rmd

To describe the general pattern in these trend lines, we regressed each state's mean contribution against a linear spline for calendar time that was estimated with five degrees of freedom. This model also contained fixed effects for state and was inversely-weighted by the standard error of each state's mean estimate within each year. We then calculated residuals for each year by substracting the meta-estimated contribution from the smoothed-observed estimated contribution and summed these residuals over time for each state to rank states according to how well the model predicted their pattern. Plots of these residuals over time were used to identify divergent states.

@@

cod_decomp_results <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/cod_decomp_results.csv")
cod_decomp_results$COD_cont_yrs_width <- cod_decomp_results$COD_cont_yrs_ucl - cod_decomp_results$COD_cont_yrs_lcl
cod_decomp_results$COD_cont_yrs_SE <- (cod_decomp_results$COD_cont_yrs_ucl - cod_decomp_results$COD_cont_yrs_lcl)/3.92
```

@@

```{r meta_regression_of_cod_contribution, echo = F}
#run 12 stratified models. bs() created the b-spline basis
models <- cod_decomp_results %>% 
  group_by(COD, sex) %>%
  do(fit = lm(COD_cont_yrs_mean ~ bs(year, degree = 1, df = 5), # + state, #remove the state FE
              weights = 1/(COD_cont_yrs_SE)^2, 
              data = .))

#calculated predicted values for all models
dfPred = broom::augment(models, fit)
dfPred$year <- 1969:2013

#merge the predictions with original dataset
cod_decomp_results <- merge(cod_decomp_results, 
                            dfPred %>% select(COD, sex, year, state, .fitted, .resid), 
                            by = c("COD", "sex", "state", "year"))

#calculated the sum of the residuals for each state-sex-cod strata (summed over years)
cod_decomp_results <- cod_decomp_results %>% 
  group_by(state, sex, COD) %>%
  mutate(sum.residual = sum(abs(.resid)))
```

@@

  a <- ggplot(subset.data,
         aes(x = year, y = COD_cont_yrs_mean)) + 
    #geom_line(aes(col = "Observed")) + 
    geom_ribbon(aes(ymin = COD_cont_yrs_lcl, ymax = COD_cont_yrs_ucl), fill = "#abd9e9") +
    geom_hline(yintercept = 0, col = "black") +
    #scale_fill_manual(values = c("#b2e2e2")) +
    facet_wrap( ~ order) + 
    geom_line(aes(y = .fitted), col = "#d7191c") +
    #scale_color_manual(values = c("#006d2c")) +
    xlab(" ") + ylab("estimated smoothed contribution vs. common trend (years)") + 
    ggtitle(paste0("Estimated smoothed contribution vs. meta-estimated common trend in ", COD1, " in ", sex1)) +
    theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1))

  b <- ggplot(subset.data,
              aes(x = year, y = .resid)) + geom_line(aes(col = "Residual")) + 
    scale_color_manual(values = c("#d7191c")) +    
    facet_wrap( ~ order) +
    xlab(" ") + ylab("Residual") +
    geom_hline(yintercept = 0, col = "black") +
      geom_text(data = subset(subset.data, year == 2010), 
                aes(y = 1, label = round(sum.residual, 1)), size = 2.5) + 
    ggtitle(paste0("Residuals vs. year in ", COD1, " in ", sex1)) +
    theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1))

    return(list(plot1 = a, plot2 = b))
}
```

@@

@@

```{r print-plot, fig.width = 9, fig.height = 9, echo = F}
ggplotly(Cardiovascular.Male.plot[[1]]) %>% layout(xaxis = list(title = "Year"))
ggplotly(Cardiovascular.Male.plot[[2]]) %>% layout(xaxis = list(title = "Year"))
```

@@

```{r print-plot-2, fig.width = 9, fig.height = 9, echo = F}
ggplotly(Cardiovascular.Female.plot[[1]]) %>% layout(xaxis = list(title = "Year"))
ggplotly(Cardiovascular.Female.plot[[2]]) %>% layout(xaxis = list(title = "Year"))
```

@@

```{r print-plot-3, fig.width = 9, fig.height = 9, echo = F}
ggplotly(Cancers.Male.plot[[1]]) %>% layout(xaxis = list(title = "Year"))
ggplotly(Cancers.Male.plot[[2]]) %>% layout(xaxis = list(title = "Year"))
```

@@

```{r print-plot-4, fig.width = 9, fig.height = 9, echo = F}
ggplotly(Cancers.Female.plot[[1]]) %>% layout(xaxis = list(title = "Year"))
ggplotly(Cancers.Female.plot[[2]]) %>% layout(xaxis = list(title = "Year"))
```

@@

```{r print-plot-5, fig.width = 9, fig.height = 9, echo = F}
ggplotly(`Non-communicable.Male.plot`[[1]]) %>% layout(xaxis = list(title = "Year"))
ggplotly(`Non-communicable.Male.plot`[[2]]) %>% layout(xaxis = list(title = "Year"))
```

@@

```{r print-plot-6, fig.width = 9, fig.height = 9, echo = F}
ggplotly(`Non-communicable.Female.plot`[[1]]) %>% layout(xaxis = list(title = "Year"))
ggplotly(`Non-communicable.Female.plot`[[2]]) %>% layout(xaxis = list(title = "Year"))
```

@@

```{r print-plot-7, fig.width = 9, fig.height = 9, echo = F}
ggplotly(Injuries.Male.plot[[1]]) %>% layout(xaxis = list(title = "Year"))
ggplotly(Injuries.Male.plot[[2]]) %>% layout(xaxis = list(title = "Year"))
```

@@

```{r print-plot-8, fig.width = 9, fig.height = 9, echo = F}
ggplotly(Injuries.Female.plot[[1]]) %>% layout(xaxis = list(title = "Year"))
ggplotly(Injuries.Female.plot[[2]]) %>% layout(xaxis = list(title = "Year"))
```

@@

```{r print-plot-9, fig.width = 9, fig.height=9, echo = F}
ggplotly(Communicable.Male.plot[[1]]) %>% layout(xaxis = list(title = "Year"))
ggplotly(Communicable.Male.plot[[2]]) %>% layout(xaxis = list(title = "Year"))
```

@@

```{r print-plot-10, fig.width = 9, fig.height=9, echo = F}
ggplotly(Communicable.Female.plot[[1]]) %>% layout(xaxis = list(title = "Year"))
ggplotly(Communicable.Female.plot[[2]]) %>% layout(xaxis = list(title = "Year"))
```

**Figure 13: Smoothed state-level trends in the contribution of all other causes to the life expectancy gap vs. the overall pattern in males**
**Figure 14: Smoothed state-level trends in the contribution of all other causes to the life expectancy gap vs. the overall pattern in males**

>>>>

To describe the general pattern in these trend lines we used random effects meta regression. Each state's mean contribution was regressed against a linear spline for calendar time that was estimated with five degrees of freedom. The model included random effects for state and was inversely-weighted by the standard error of each state's mean estimate within each year. To identify divergent states, we compared the state-specific trends (estimated separately for each state using the Bayesian model) to the general pattern estimated using meta regression.

@@

cod_decomp_results <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/cod_decomp_results.csv")
cod_decomp_results$COD_cont_yrs_width <- cod_decomp_results$COD_cont_yrs_ucl - cod_decomp_results$COD_cont_yrs_lcl
cod_decomp_results$COD_cont_yrs_SE <- (cod_decomp_results$COD_cont_yrs_ucl - cod_decomp_results$COD_cont_yrs_lcl)/3.92

cod_decomp_results <- reorder.as.map(cod_decomp_results, "state")
```

@@

```{r meta_regression_of_cod_contribution, echo = F}
#model option 1: FE analysis

#run 12 stratified models. bs() created the b-spline basis
models <- cod_decomp_results %>% 
  group_by(COD, sex) %>%
  do(fit = lm(COD_cont_yrs_mean ~ bs(year, degree = 1, df = 5) + state,
              weights = 1/(COD_cont_yrs_SE)^2, 
              data = .))

#calculated predicted values for all models
dfPred <- broom::augment(models, fit)
dfPred$year <- 1969:2013

#merge the predictions with original dataset
cod_decomp_results <- merge(cod_decomp_results, 
                            dfPred %>% select(COD, sex, year, state, .fitted, .resid), 
                            by = c("COD", "sex", "state", "year")) %>%
                      rename(fitted.FE = .fitted, resid.FE = .resid)

#calculated the sum of the residuals for each state-sex-cod strata (summed over years)
cod_decomp_results <- cod_decomp_results %>% 
  group_by(state, sex, COD) %>%
  mutate(sum.residual = sum(abs(resid.FE)))
```

```{r meta_RE_analysis, echo = F}
library(lme4)

#run 12 stratified models. bs() created the b-spline basis
RE.models <- cod_decomp_results %>% 
  group_by(COD, sex) %>%
  do(fit = lmer(COD_cont_yrs_mean ~ bs(year, degree = 1, df = 5) + (1|state),
              weights = 1/(COD_cont_yrs_SE)^2, 
              data = .))

#calculated predicted values for all models
RE.pred <- broom::augment(RE.models, fit)
RE.pred$year <- 1969:2013

RE.models$RE.estimates <- purrr::map(RE.models$fit, ~ ranef(.x))

sex <- rep(c(rep("Female", 40), rep("Male", 40)), 6)
COD <- c(rep("All other causes", 80), rep("Cancers", 80), rep("Cardiovascular", 80),
         rep("Communicable", 80), rep("Injuries", 80), rep("Non-communicable", 80))
RE.estimates <- c(RE.models$RE.estimates[[1]]$state$`(Intercept)`, RE.models$RE.estimates[[2]]$state$`(Intercept)`, 
                  RE.models$RE.estimates[[3]]$state$`(Intercept)`, RE.models$RE.estimates[[4]]$state$`(Intercept)`, 
                  RE.models$RE.estimates[[5]]$state$`(Intercept)`, RE.models$RE.estimates[[6]]$state$`(Intercept)`, 
                  RE.models$RE.estimates[[7]]$state$`(Intercept)`, RE.models$RE.estimates[[8]]$state$`(Intercept)`, 
                  RE.models$RE.estimates[[9]]$state$`(Intercept)`, RE.models$RE.estimates[[10]]$state$`(Intercept)`,
                  RE.models$RE.estimates[[11]]$state$`(Intercept)`, RE.models$RE.estimates[[12]]$state$`(Intercept)`)
state <- rep(levels(cod_decomp_results$state), 12)
REs <- data.frame(sex, COD, RE.estimates, state)

cod_decomp_results <- merge(cod_decomp_results,
                            REs, by = c("COD", "sex", "state"),
                            all.x = T)

#merge the predictions with original dataset
cod_decomp_results <- merge(cod_decomp_results, 
                             RE.pred %>% select(COD, sex, state, year, 
                                                bs.year..degree...1..df...5.,
                                                .fitted, .resid, .fixed, .mu), 
                             by = c("COD", "sex", "state", "year"))

cod_decomp_results <- cod_decomp_results %>% rename(fitted.RE = .fitted, fixed.from.RE = .fixed)

# test1 <- lmer(COD_cont_yrs_mean ~ bs(year, degree = 1, df = 5) + (1|state), weights = 1/(COD_cont_yrs_SE)^2, data = subset(cod_decomp_results, COD == "Cardiovascular" & sex == "Male"))
# 
# pred.general.trend <- predict(test1, re.form = NA)
# pred.with.REs <- predict(test1)
# 
# plot(y=pred.general.trend[1:45], x=1969:2013)
```

@@

  a <- ggplot(subset.data,
         aes(x = year, y = COD_cont_yrs_mean)) + 
    geom_hline(aes(yintercept = 0)) + geom_vline(aes(xintercept = 1969)) + #geom_hline(yintercept = 0, col = "black") +
    #geom_line(aes(col = "Observed")) + 
    geom_ribbon(aes(ymin = COD_cont_yrs_lcl, ymax = COD_cont_yrs_ucl), fill = "#abd9e9") +
    #geom_line(aes(col = Census_Division)) + #add this for the online version, and then adjust legend and hover display    
    #scale_fill_manual(values = c("#b2e2e2")) +
    facet_wrap( ~ state.map.order, ncol = 11, drop = F) + 
    geom_line(aes(y = fitted.FE), col = "#d7191c") +
    ggtitle(paste0("Estimated smoothed contribution vs. meta-estimated common trend in ", COD1, " in ", sex1)) +
    theme_classic(base_size = 10) + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          strip.background=element_blank(),
        axis.line=element_blank(),
        axis.ticks=element_blank()) +
    ylab("estimated smoothed contribution vs. common trend (years)") +
    xlab("Year (1969-2013)")

  b <- ggplot(subset.data,
              aes(x = year, y = resid.FE)) + geom_line(aes(col = "Residual")) + 
    scale_color_manual(values = c("#d7191c")) +    
    facet_wrap( ~ order) +
    xlab(" ") + ylab("Residual") +
    geom_hline(yintercept = 0, col = "black") +
      geom_text(data = subset(subset.data, year == 2010), 
                aes(y = 1, label = round(sum.residual, 1)), size = 2.5) + 
    ggtitle(paste0("Residuals vs. year in ", COD1, " in ", sex1)) +
    theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1))

  c <- ggplot(subset.data,
              aes(x = year, y = COD_cont_yrs_mean)) + 
    geom_hline(aes(yintercept = 0)) + geom_vline(aes(xintercept = 1969)) + #geom_hline(yintercept = 0, col = "black") +

    #geom_line(aes(col = "Observed")) + 
    geom_ribbon(aes(ymin = COD_cont_yrs_lcl, ymax = COD_cont_yrs_ucl), fill = "#abd9e9") +
    #geom_line(aes(col = Census_Division)) + #add this for the online version, and then adjust legend and hover display    
    #scale_fill_manual(values = c("#b2e2e2")) +
    facet_wrap( ~ state.map.order, ncol = 11, drop = F) + 
    geom_line(aes(y = fitted.RE - RE.estimates), col = "#d7191c") +
    #geom_line(aes(y = fitted.RE), col = "#d7191c", lty = 3) +
    ggtitle(paste0("Estimated smoothed contribution vs. meta-estimated common trend in ", COD1, " in ", sex1)) +
    theme_classic(base_size = 10) + 
    theme(axis.text.x = element_blank(),
          strip.background=element_blank(),
          axis.line=element_blank(),
          axis.ticks=element_blank()) +
    ylab("estimated smoothed contribution vs. common trend (years)") +
    xlab("Year (1969-2013)")

  return(list(plot1 = a, plot2 = b, plot3 = c))
}
```

@@

```{r refine_plots, eval = F, echo = F}
#need tp add the "grid lines post-hoc" because they depend on the max height of the plot

    geom_hline(aes(yintercept = 1), lwd = 0.5, col = "grey") +
    geom_hline(aes(yintercept = 2), lwd = 0.5, col = "grey") +
    geom_hline(aes(yintercept = 3), lwd = 0.5, col = "grey") +
    geom_hline(aes(yintercept = 4), lwd = 0.5, col = "grey") +

#turn hover off for grid lines and axis lines
refine.patterns <- function(plot) {
  for(i in 1:320){
  if(plot$x$data[[i]]$line$color == "rgba(190,190,190,1)"){
    plot$x$data[[i]]$line$width <- 0.7559055
    plot$x$data[[i]]$hoverinfo <- "none"
  }
    if(plot$x$data[[i]]$line$color == "rgba(0, 0, 0, 1)"){
    plot$x$data[[i]]$hoverinfo <- "none"
  }
}
}

```

@@

```{r print-plot, fig.width = 9, fig.height = 9, echo = F}
ggplotly(Cardiovascular.Male.plot[[3]]) %>% layout(xaxis = list(title = "Year"))
```

@@

```{r print-plot-2, fig.width = 9, fig.height = 9, echo = F}
#ggplotly(Cardiovascular.Female.plot[[1]]) %>% layout(xaxis = list(title = "Year"))
ggplotly(Cardiovascular.Female.plot[[3]]) %>% layout(xaxis = list(title = "Year"))
```

@@

```{r print-plot-3, fig.width = 9, fig.height = 9, echo = F}
#ggplotly(Cancers.Male.plot[[1]]) %>% layout(xaxis = list(title = "Year"))
ggplotly(Cancers.Male.plot[[3]]) %>% layout(xaxis = list(title = "Year"))
```

@@

```{r print-plot-4, fig.width = 9, fig.height = 9, echo = F}
#ggplotly(Cancers.Female.plot[[1]]) %>% layout(xaxis = list(title = "Year"))
ggplotly(Cancers.Female.plot[[3]]) %>% layout(xaxis = list(title = "Year"))
```

@@

```{r print-plot-5, fig.width = 9, fig.height = 9, echo = F}
#ggplotly(`Non-communicable.Male.plot`[[1]]) %>% layout(xaxis = list(title = "Year"))
ggplotly(`Non-communicable.Male.plot`[[3]]) %>% layout(xaxis = list(title = "Year"))
```

@@

```{r print-plot-6, fig.width = 9, fig.height = 9, echo = F}
#ggplotly(`Non-communicable.Female.plot`[[1]]) %>% layout(xaxis = list(title = "Year"))
ggplotly(`Non-communicable.Female.plot`[[3]]) %>% layout(xaxis = list(title = "Year"))
```

@@

```{r print-plot-7, fig.width = 9, fig.height = 9, echo = F}
#ggplotly(Injuries.Male.plot[[1]]) %>% layout(xaxis = list(title = "Year"))
ggplotly(Injuries.Male.plot[[3]]) %>% layout(xaxis = list(title = "Year"))
```

@@

```{r print-plot-8, fig.width = 9, fig.height = 9, echo = F}
#ggplotly(Injuries.Female.plot[[1]]) %>% layout(xaxis = list(title = "Year"))
ggplotly(Injuries.Female.plot[[3]]) %>% layout(xaxis = list(title = "Year"))
```

@@

```{r print-plot-9, fig.width = 9, fig.height=9, echo = F}
#ggplotly(Communicable.Male.plot[[1]]) %>% layout(xaxis = list(title = "Year"))
ggplotly(Communicable.Male.plot[[3]]) %>% layout(xaxis = list(title = "Year"))
```

@@

```{r print-plot-10, fig.width = 9, fig.height=9, echo = F}
#ggplotly(Communicable.Female.plot[[1]]) %>% layout(xaxis = list(title = "Year"))
ggplotly(Communicable.Female.plot[[3]]) %>% layout(xaxis = list(title = "Year"))
```

**Figure 13: Smoothed state-level trends in the contribution of all other causes to the life expectancy gap vs. the overall pattern in males**

```{r print-plot-11, fig.width = 9, fig.height=9, echo = F}
#ggplotly(Communicable.Female.plot[[1]]) %>% layout(xaxis = list(title = "Year"))
ggplotly(`All other causes.Male.plot`[[3]]) %>% layout(xaxis = list(title = "Year"))
```

**Figure 14: Smoothed state-level trends in the contribution of all other causes to the life expectancy gap vs. the overall pattern in males**

```{r print-plot-12, fig.width = 9, fig.height=9, echo = F}
#ggplotly(Communicable.Female.plot[[1]]) %>% layout(xaxis = list(title = "Year"))
ggplotly(`All other causes.Female.plot`[[3]]) %>% layout(xaxis = list(title = "Year"))
```
